brados: declarative, programmable object storage

### 2016-05-25 paper.tex
updates
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that future proofs an
implementation against hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knoweldge
addressed portability concerns that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

@@

As an example, one valid design option is to store each log position in an
object with a name using a one-to-one mapping with the log entry position.
This would simplify the design of the \emph{write} interface because a small
amount of metadata stored as a header in the object could describe the state
of the log entry.  However, as we will see this choice of a physical design
can result in poor performance compared to other designs. In the remainder of
this section we will define the entire design space and use a set of targetted
benchmarks to arrive at a final design. Finally we will show that these design
decisions can lead to non-optimal decisions and suggest an automated approach
is desirable.

@@

\textbf{Ceph Versions}: to show the evolution of Ceph's performance and
behavior, we ran the same benchmark, which measures append throughput, with the
same configurations, hardware, and tunables on 2 versions of Ceph: Jewel (April
2016) - the newest stable release and Firefly (May 2014) - a long term support
version that introduced cache tiering and erasure coding. The Jewel row of 
Figure~\ref{fig:jewel_v_firefly_v_es} shows that the size of the log entries
has a clear impact on performance but the Firefly release shows no such trend.
The fact that the appends per second are an order of magnitude slower on
Firefly indicates that the the benchmark is bottlenecked by the Ceph software
instead of the external log implementation.

The bytestream interface outperforms the key-value interface by 12\%, compared
to over 100\% in the newer version of Ceph. Qualitatively, given the reduced
overall throughput achieved in the older version, some developers may find that
incurring the overhead of using the key-value interface is an easy decision
given the reduced complexity of the design space for metadata management. If
this choice had been made, then a future system upgrade could drop a
significant number of iops on the floor.

\textbf{Ceph Features}: to show the affect of cutting-edge Ceph features, we
now show results for BlueStore~\cite{weil:vault2016-bluestore}. BlueStore is a
replacement for the FileStore file system in the OSD (traditionally XFS).
FileStore has performance problems with transactions and enumerations; namely
the journal needed to assure atomicity incurs double writes and the file system
metadata model makes object listings slow, respectively. BlueStore stores data
directly on a block device and the metadata in RocksDB, which is provided by a
minimalistic, non-POSIX C++ filesystem. This model adheres to the overall
software defined storage strategy of Ceph because it gives the administrator
the flexiblity to store the 3 componenets of BlueStore (e.g., data, RocksDB
database, and RocksDB write-ahead log) on any partition on any device in the
OSD.

\textbf{Takeaway}: choosing the best implementations is dependent on both the
timing of the developemnt (Ceph Version) and the expertise of the administrator
(Ceph Features). Different versions and features of Ceph may lead the
administrator to choose a suboptimal implementation for the system's next
upgrade. The software parameters must be accounted for and benchmarked when
making design decisions.

@@

The most recent version of Ceph (v10.2.0-1281-g1f03205) has 994 tunable
parameters\footnote{This number comes from \texttt{src/common/config\_opts.h}
with debug options filtered out.}, where 195 of them pertain to the OSD itself
and 95 of them focus on the OSD back end file system (i.e. its
\texttt{filestore}). Ceph also has tunables for the subystems it uses, like
LevelDB (10 tunables), RocksDB (5 tunables), its own key-value stores (5
tunables), its object cache (6 tunables), its journals (24 tunables), and its
other optional filestores like BlueStore (49 tunables).

This many domain-specific tunables makes it almost impossible to come up with
the best set of tunables, although auto-tuning like the work done
in~\cite{behzad:sc2013-autotuning} could go a long way. Regardless of the
technique that we use, it is clear the the number of tunables increases the
physical design parameters to an unwieldy state space size.

@@

This physical design process, of first thinking about the design trade-offs and
then doing parameters sweeps, is insufficient for large general-purpose systems
like Ceph. Had we built CORFU from the ground up the story might be
different but to deal with the large state space we need a way to automatically
and autonomically decide which parameters to choose.

@@

As Ceph continues to develop into a target for building distributed services
and applications, it is important to not restrict use to developers seasonsed
in distributed programming and those with knowledge of the intracacies of Ceph
and its performance model. The hard-coded imperative method of writing object
interfaces today restrict the optimizations that can be performed
automatically or derived from static analysis.  To this end we present Brados,
a declarative language with relational semantics for defining the behavior of
RADOS object interfaces.  Brados uses a declarative data model, can reduce the
learning curve for new users, and allows existing developers to increase
productivity by writing less code that is more portable.

@@

Trimming a log entry always succeeds. In Listing~\ref{lst:trim} the
\texttt{<+-} operator simulatenously removes the log entry with the given 
position and replaces it with an entry with its state set to \emph{trimmed}.
In practice the removal of a log entry may trigger garbage collection, but
we model it here as an update for brevity.

@@

The write and fill interfaces are implemented similarily, and are both shown in
Listing~\ref{lst:write}.  A \texttt{valid\_write} collection is created if the
operation position is not found in the log.  A \texttt{valid\_write} operation
is then merged into log, otherwise a read only error is returend indicating
that the log position was already written to.  The fill operation is identical
except a the fill state is set on the log entry.

@@

Finally the read interface is shown in Listing~\ref{lst:read}, and structured
in a similar way to the write and fill interfaces. First we create a collection
containing a valid read operation (named \texttt{ok\_read}) that is in the log
and does not have the filled or trimmed state set.  The data read from the log
is returend in the case of a valid read operation, otherwise an error is
returend through the output interface.

@@

But beyond the convience of writing less code, it is far easier for the
programmer writing an interface such as CORFU to convenience herself of the
correctness of the high-level details of the implementation without being
distracted by issues related to physical design or the many other gotchas that
one must deal with when writing low-level systems software.

@@

The challenge of navigating the physical design space has served as the
primary source of motivation for selection of a declarative langauge. While
our implementation does not yet

The Bloom language that we use as a basis for Brados produces a data flow
graph that can be used in static analysis. We envision that this graph will be
made available to the OSD and used to reorder and collesce requests based on
optimization criteria available from a performance model combined with
semantic information from the dataflow. For example today object classes are
represented as black boxes from the point of view of the OSD execution engine.
Understanding the behavior of an object class may allow intelligent
prefetching. Another type of analysis that may be useful for optimization is
optimistic execution combined with branch prediction where frequent paths
through a dataflow are handled optimistically.

@@

{\bf Active storage.}
There is a wide variety of work related to \emph{active storage} that seeks to
increase performance and reduce data movement by exploiting remote storage
resources. This concept has been applied in the context of on-disk
controllers and object-based storage
(T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
concerns have examined using managed lanuages, sandbox technologies, as well
as restricting extension installation to
vendors~\cite{john:hiperio08,xie:msst11,runde:msst12}.

@@

Most closely related to our work are efforts that consider specific
programming models in the context of active storage such as a stream-based
model~\cite{acharya:asplos98}, a model that assisted in optmiizing balancing
computation~\cite{wickremesinghe:hpdc022}, and optimization that took
advantage of read-only data to reorder operations~\cite{huston:fast04}.
While our goals are similar to previous work, we start with a declarative
language that will enable us to apply optimization techniques beyond
the limited domain-specific optimizations found in previous work.

>>>>

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that future proofs an
implementation against hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knowledge
addressed portability concerns that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

@@

As an example, one valid design option is to store each log position in an
object with a name using a one-to-one mapping with the log entry position.
This would simplify the design of the \emph{write} interface because a small
amount of metadata stored as a header in the object could describe the state
of the log entry.  However, as we will see this choice of a physical design
can result in poor performance compared to other designs. In the remainder of
this section we will define the entire design space and use a set of targeted
benchmarks to arrive at a final design. Finally we will show that these design
decisions can lead to non-optimal decisions and suggest an automated approach
is desirable.

@@

\textbf{Ceph Versions}: to show the evolution of Ceph's performance and
behavior, we ran the same benchmark, which measures append throughput, with the
same configurations, hardware, and tunables on 2 versions of Ceph: Jewel (April
2016) - the newest stable release and Firefly (May 2014) - a long term support
version that introduced cache tiering and erasure coding. The Jewel row of 
Figure~\ref{fig:jewel_v_firefly_v_es} shows that the size of the log entries
has a clear impact on performance but the Firefly release shows no such trend.
The fact that the appends per second are an order of magnitude slower on
Firefly indicates that the benchmark is bottlenecked by the Ceph software
instead of the external log implementation.

The bytestream interface outperforms the key-value interface by 12\%, compared
to over 100\% in the newer version of Ceph. Qualitatively, given the reduced
overall throughput achieved in the older version, some developers may find that
incurring the overhead of using the key-value interface is an easy decision
given the reduced complexity of the design space for metadata management. If
this choice had been made, then a future system upgrade could drop a
significant number of IOPS on the floor.

\textbf{Ceph Features}: to show the affect of cutting-edge Ceph features, we
now show results for BlueStore~\cite{weil:vault2016-bluestore}. BlueStore is a
replacement for the FileStore file system in the OSD (traditionally XFS).
FileStore has performance problems with transactions and enumerations; namely
the journal needed to assure atomicity incurs double writes and the file system
metadata model makes object listings slow, respectively. BlueStore stores data
directly on a block device and the metadata in RocksDB, which is provided by a
minimalistic, non-POSIX C++ filesystem. This model adheres to the overall
software defined storage strategy of Ceph because it gives the administrator
the flexibility to store the 3 components of BlueStore (e.g., data, RocksDB
database, and RocksDB write-ahead log) on any partition on any device in the
OSD.

\textbf{Takeaway}: choosing the best implementations is dependent on both the
timing of the development (Ceph Version) and the expertise of the administrator
(Ceph Features). Different versions and features of Ceph may lead the
administrator to choose a suboptimal implementation for the system's next
upgrade. The software parameters must be accounted for and benchmarked when
making design decisions.

@@

The most recent version of Ceph (v10.2.0-1281-g1f03205) has 994 tunable
parameters\footnote{This number comes from \texttt{src/common/config\_opts.h}
with debug options filtered out.}, where 195 of them pertain to the OSD itself
and 95 of them focus on the OSD back end file system (i.e. its
\texttt{filestore}). Ceph also has tunables for the subsystems it uses, like
LevelDB (10 tunables), RocksDB (5 tunables), its own key-value stores (5
tunables), its object cache (6 tunables), its journals (24 tunables), and its
other optional object stores like BlueStore (49 tunables).

This many domain-specific tunables makes it almost impossible to come up with
the best set of tunables, although auto-tuning like the work done
in~\cite{behzad:sc2013-autotuning} could go a long way. Regardless of the
technique that we use, it is clear the number of tunables increases the
physical design parameters to an unwieldy state space size.

@@

This physical design process, of first thinking about the design trade-offs and
then doing parameters sweeps, is insufficient for large general-purpose systems
like Ceph. Had we built CORFU from the ground up the story might be
different but to deal with the large state space we need a way to automatically
decide which parameters to choose.

@@

As Ceph continues to develop into a target for building distributed services
and applications, it is important to not restrict use to developers seasoned
in distributed programming and those with knowledge of the intricacies of Ceph
and its performance model. The hard-coded imperative method of writing object
interfaces today restrict the optimizations that can be performed
automatically or derived from static analysis.  To this end we present Brados,
a declarative language with relational semantics for defining the behavior of
RADOS object interfaces.  Brados uses a declarative data model, can reduce the
learning curve for new users, and allows existing developers to increase
productivity by writing less code that is more portable.

@@

Trimming a log entry always succeeds. In Listing~\ref{lst:trim} the
\texttt{<+-} operator simultaneously removes the log entry with the given 
position and replaces it with an entry with its state set to \emph{trimmed}.
In practice the removal of a log entry may trigger garbage collection, but
we model it here as an update for brevity.

@@

The write and fill interfaces are implemented similarly, and are both shown in
Listing~\ref{lst:write}.  A \texttt{valid\_write} collection is created if the
operation position is not found in the log.  A \texttt{valid\_write} operation
is then merged into log, otherwise a read only error is returned indicating
that the log position was already written to.  The fill operation is identical
except the fill state is set on the log entry.

@@

Finally the read interface is shown in Listing~\ref{lst:read}, and structured
in a similar way to the write and fill interfaces. First we create a collection
containing a valid read operation (named \texttt{ok\_read}) that is in the log
and does not have the filled or trimmed state set.  The data read from the log
is returned in the case of a valid read operation, otherwise an error is
returned through the output interface.

@@

But beyond the convenience of writing less code, it is far easier for the
programmer writing an interface such as CORFU to convenience herself of the
correctness of the high-level details of the implementation without being
distracted by issues related to physical design or the many other gotchas that
one must deal with when writing low-level systems software.

@@

The challenge of navigating the physical design space has served as the
primary source of motivation for selection of a declarative language. While
our implementation does not yet

The Bloom language that we use as a basis for Brados produces a data flow
graph that can be used in static analysis. We envision that this graph will be
made available to the OSD and used to reorder and coalesce requests based on
optimization criteria available from a performance model combined with
semantic information from the dataflow. For example today object classes are
represented as black boxes from the point of view of the OSD execution engine.
Understanding the behavior of an object class may allow intelligent
prefetching. Another type of analysis that may be useful for optimization is
optimistic execution combined with branch prediction where frequent paths
through a dataflow are handled optimistically.

@@

{\bf Active storage.}
There is a wide variety of work related to \emph{active storage} that seeks to
increase performance and reduce data movement by exploiting remote storage
resources. This concept has been applied in the context of on-disk
controllers and object-based storage
(T10)~\cite{riedel:vldb98,du:nwesp05,xie:msst11}.  Solutions to safety
concerns have examined using managed languages, sandbox technologies, as well
as restricting extension installation to
vendors~\cite{john:hiperio08,xie:msst11,runde:msst12}.

@@

Most closely related to our work are efforts that consider specific
programming models in the context of active storage such as a stream-based
model~\cite{acharya:asplos98}, a model that assisted in optimizing balancing
computation~\cite{wickremesinghe:hpdc022}, and optimization that took
advantage of read-only data to reorder operations~\cite{huston:fast04}.
While our goals are similar to previous work, we start with a declarative
language that will enable us to apply optimization techniques beyond
the limited domain-specific optimizations found in previous work.
